CDSR Human Data/Modelling Paper

### 2013-09-11 scenarios/cdsr/trunk/papers/ACS2013/main.tex
changed bits to related work

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points. 
%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.

@@

\section{Modeling - Nick}

As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g. understanding spatially-situated task utterances), they must have access to internal models which can represent CDSRs and recognise them in new environments. Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to work. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain code for image region representations~\cite{Freeman:1961}. This is an approach we followed in previous work, where boundary segments connected vertices defined by contextually-appropriate anchor points~\cite{Hawes:2012}.  Whilst our previous approach provided basic performance, the crisp boundaries it generated caused mismatches with human data over the extent of regions. Methods based on underlying spatial functions may be more flexible in this regard, as they can be selectively thresholded depending on context.

% Basic approaches

\textbf{TODO:} Perhaps add related work on robot spatial representations, e.g. maps, rooms.

In the following sections we explore whether two possible CDSR models (one based on boundary segments, one based on a spatial function) are able to satisfy the previously described desiderata for a cognitive systems approach to representing and recognising CDSRs. The surrounding discussion is intended to explore the strengths and weaknesses of these models, providing directions for future research.

\subsection{Symbolic Boundary Method - KLENK}
In our previous work \cite{Hawes:2012}, we presented an approach on a mobile robot that combined qualitative spatial relations, object recognition, and analogy to transfer CDSRs represented as polygons defined by \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11}.  Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).  An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.  In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.  Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.  In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.

@@

This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task.

Applying this model to the points and regions indicated in our experiment illustrates two limitations.  Regions do not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way into the corners.  In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.  This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.

\subsection{Spatial Function Method - NICK}

Overview - either Chris's approach or something similar
Example
Results?
Discussion 
Where do the points come from

@@

\begin{acknowledgements} 
\noindent
STRANDS etc.
\end{acknowledgements}

>>>>

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sample t-tests and in all cases the ranking of the sweetspot was found to be significantly higher than either of the other two points using a confidence level of $95\%$. 
%There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli. Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j). This higher rating is statiscially significant at a $95\%$ confidence level. Stimuli (k) and (l) present and analagous situation with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

\section{Related Work}

As the results above show, in order for an artificial cognitive system to operate successfully with humans (e.g. understanding spatially-situated task utterances), they must have access to internal models which can both represent CDSRs and recognise them in new environments. Whilst spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}), or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines), as in chain codes for image region representations~\cite{Freeman:1961}.

This is an approach we followed in previous work \cite{Hawes:2012}, where we developed an approach that defined CDSRs using \textit{anchor points} \cite{DBLP:journals/jetai/KlenkFTK11} as the vertices of the region polygon. Anchor points are symbolic descriptions that link a conceptual entity (e.g., the front of a classroom) to perceived entities (e.g., the lower left extent of a desk).  Once a particular representation has been modelled in this manner, we were able to find the same type of region in previously unseen rooms using analogical transfer. An example of the representation and transfer approach is Figure \ref{fig:anchor-point-transfer}.  In the know example, a human defines the front of the classroom as a polygon consisting of four anchor points.  Analogical transfer works by aligning structured representations and transferring partially aligned statements from the known example to the new environment.  In this case, the anchor points are transferred from the group of desks in the known example to the corresponding group of desks in the new environment.

@@

This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task. Applying this model to the points and regions indicated in our experiment illustrates two limitations.  Regions do not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way into the corners.  In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.  This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.

The use of spatial regions connected to function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Such fields can only vary with respect to context in terms of their size, which does not make them particularly well suited to modelling CDSRs.

@@

\begin{acknowledgements} 
\noindent
The research leading to these results has received funding from the European Union Seventh Framework Program (FP7/2007-2013) under the project STRANDS (grant agreement No 600623), and from the EPSRC under grant EP/K014293/1.
\end{acknowledgements}
